\chapter*{Abstract}
\label{cha:abstract}

This master thesis presents a novel web extension designed to bolster user privacy by effectively blocking tracking
requests in real-time. Leveraging machine learning models, the extension demonstrates remarkable performance in identifying
and intercepting web tracking requests. Two models, namely Model 1 and Model 2, were evaluated using the Tranco Top 1K websites dataset.
Model 1 achieved a precision of 0.98 and a recall of 0.99 for non-trackers, with a precision of 0.97, a recall of 0.96, and a
F1-score of 0.96 for trackers. Similarly, Model 2 exhibited outstanding precision, recall, and F1-scores of 0.99, 0.98, and 0.98
for non-trackers, and 0.95, 0.97, and 0.96 for trackers, respectively.

The real-time aspect of the web extension ensures immediate action upon the detection of tracking requests. By intercepting
and blocking trackers in real-time, user privacy is significantly enhanced, minimizing the risks associated with data collection
and online tracking. This dynamic approach empowers the use of ML models in the web tracking enviroment which has been dominated 
by manually curated blocking lists. The performance of the models in a real-time enviroment shows results close to the Brave browsers'
advertisment and tracking blocker.

To facilitate seamless deployment and continuous improvement, a robust backend infrastructure has been developed. This infrastructure
enables effortless integration of new machine learning models into the web extension, ensuring adaptability to evolving tracking
techniques and enhancing the extension's efficacy over time. By leveraging this backend infrastructure, the web extension remains
up-to-date and can effectively counter emerging privacy threats, providing users with ongoing protection.

Through rigorous evaluation and testing, the web extension has demonstrated its ability to block tracking requests effectively.
The incorporation of machine learning models enables accurate identification and interception of trackers. 
The real-time nature of the extension ensures immediate action, preventing unauthorized data
collection and bolstering user privacy. With a scalable backend infrastructure in place, the web extension remains agile and
adaptable, poised to counter evolving tracking methods. This master thesis contributes to the advancement of privacy protection
in the digital landscape, offering users a ML based solution to combat online tracking and safeguard their personal information.


